Fusion of Mass Spectrometry and Raman Spectroscopy Data

### 2024-06-06 dev/dev_paracetamol.Rmd
Updated statistic engine with classification

fusedMat <- cbind(mmSpec, mrSpec)

attr(fusedMat, "xaxis.name") = "Keys"
attr(fusedMat, "xaxis.values") = seq_len(ncol(fusedMat))

# Mean centering improves separation but can we do it?
library(mdatools)
fusedMat = prep.autoscale(fusedMat, center = TRUE, scale = FALSE)
```

@@

## Overall PCA

```{r, include=FALSE}
library(mdatools)
m = pca(fusedMat, 3, center = FALSE, scale = FALSE, info = "Overall PCA model")
```

```{r ,echo=FALSE,fig.width=10,fig.height=4}
library(plotly)

cl <- StreamFind:::.get_colors(rownames(fusedMat))
scores_polt <- plot_ly()

scores_polt <- scores_polt %>% add_trace(
  x = m$res$cal$scores[, "Comp 1"],
  y = m$res$cal$scores[, "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  color = names(cl),
  colors = cl,
  text = rownames(m$res$cal$scores),
  textfont = list(size = 14, color = cl),
  textposition = "top",
  marker = list(size = 10, color = cl),
  showlegend = FALSE
)

scores_polt <- scores_polt %>% layout(
  xaxis = list(title = "Principal Component 1"),
  yaxis = list(title = "Principal Component 2")
)

cl2 <- StreamFind:::.get_colors(1:2)
loadings_plot <- plot_ly()

loadings_plot <- loadings_plot %>% add_trace(
  x = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 1"],
  y = m$loadings[rownames(m$loadings) %in% colnames(mrSpec), "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mrSpec))],
  textfont = list(size = 14, color = cl2[1]),
  textposition = "top",
  legendgroup = "Raman",
  name = "Raman",
  marker = list(size = 10, color = cl2[1])
)

loadings_plot <- loadings_plot %>% add_trace(
  x = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 1"],
  y = m$loadings[rownames(m$loadings) %in% colnames(mmSpec), "Comp 2"],
  type = "scatter",
  mode = "markers+text",
  text = rownames(m$loadings)[(rownames(m$loadings) %in% colnames(mmSpec))],
  textfont = list(size = 14, color = cl2[2]),
  textposition = "top",
  legendgroup = "MS",
  name = "MS",
  marker = list(size = 10, color = cl2[2])
)

loadings_plot <- loadings_plot %>% layout(
  xaxis = list(title = "Principal Component 1"),
  yaxis = list(title = "Principal Component 2")
)

subplot(
  scores_polt, loadings_plot, nrows = 1,
  shareX = FALSE, shareY = FALSE, titleX = TRUE, titleY = TRUE,
  margin = 0.1
)
```

@@

```{r}
# rownames(fusedMat)
main_set <- fusedMat[c(14:15, 18), ]
test_set <- fusedMat[c(1:13, 16:17, 19:30), ]

m3 = pca(main_set, 2, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 1E-1, gamma = 1E-2)

plot(m3)

res = predict(m3, test_set)

c = categorize(m3, res)

names(c) <- rownames(test_set)

plotResiduals(m3, 
  res = list("train" = m3$res$cal, "new" = res),
  norm = TRUE, show.labels = TRUE,
  lim.col = c("red", "orange"), lim.lwd = c(2, 2), lim.lty = c(1, 2)
)

print(c)
```

@@

```{r}
train_set <- fusedMat[c(grep("_1", rownames(fusedMat)), grep("_2", rownames(fusedMat))), ]

train_set <- train_set[order(rownames(train_set)), ]

@@

labels <- factor(c(rep("ass", 2), rep("caff", 2), rep("fiz_acid", 2), rep("fiz_ass", 2), rep("fiz_caff", 2),
  rep("fiz_altered", 2), rep("fiz_altered", 2), rep("fiz", 2), rep("fiz", 2), rep("par", 2))
)

library(class)

knn_model <- knn(train = train_set, test = test_set, cl = labels, k = 3, prob = FALSE)

names(knn_model) <- rownames(test_set)

data.frame(class = knn_model)
# print(knn_model)
```

@@

>>>>

fusedMat <- cbind(mmSpec, mrSpec)
```

@@

@@

```{r}
main_set <- fusedMat[c(8:9, 10, 12, 14:15, 17:18), ]
test_set <- fusedMat[c(1:7, 11, 13, 16, 19:30), ]

stat <- StatisticEngine$new(data = main_set)
stat$make_pca_model(Settings_make_pca_model_mdatools(ncomp = 3, alpha = 0.005, gamma = 0.0005))
stat$predict(test_set)
stat$plot_predicted_distances()
```

@@

```{r}
train_set <- fusedMat[c(grep("_1", rownames(fusedMat)), grep("_2", rownames(fusedMat))), ]
train_set <- train_set[order(rownames(train_set)), ]

@@

labels <- c(
  rep("ass", 2), rep("caff", 2), rep("fiz_acid", 2), rep("fiz_ass", 2), rep("fiz_caff", 2),
  rep("fiz_altered", 2), rep("fiz_altered", 2), rep("fiz", 2), rep("fiz", 2), rep("par", 2)
)

stat2 <- StatisticEngine$new(data = train_set)
stat2$add_classes(labels)
stat2$prepare_classification(Settings_prepare_classification_knn(k = 3))
stat2$classify(test_set)
stat2$classification_results
```

@@

## Old code

```{r include=FALSE}
# library(mdatools)
# m3 = pca(main_set, 3, center = FALSE, scale = FALSE, info = "Unsupervised PCA model", alpha = 0.005, gamma = 0.0005)
# plot(m3, show.labels = TRUE)
# res = predict(m3, test_set)
# plot(res, show.labels = TRUE)
# c = categorize(m3, res, 3)
# names(c) <- rownames(test_set)
# c
# plotResiduals(m3,
#   ncomp = 3,
#   res = list("train" = m3$res$cal, "new" = res),
#   norm = TRUE, show.labels = TRUE,
#   lim.col = c("red", "orange"), lim.lwd = c(2, 2), lim.lty = c(1, 2)
# )
# plotResiduals(ncomp = 1, res, show.plot = FALSE, norm = TRUE)
# print(c)
# is(m3$res$cal, "ldecomp")
```
